---
title: "Homework 1" 
subtitle: "PSTAT 231"
author: "Anna Kazachka"
output: html_document
date: "March 3, 2021"
---
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(tidymodels)
library(tidyselect)
library(ggplot2)
library(graphics)
library(base)
```

# Machine Learning Main Ideas
## Question 1: 
#### Define supervised and unsupervised learning. What are the difference(s) between them?
Supervised learning maps an input to an output using already existing pairs. Unsupervised learning models unlabeled data by discovering patterns and finding the structure of the inputs. Unsupervised learning is left to work through the data on its own and is not given labeled data while supervised learning is. 

## Question 2:
#### Explain the difference between a regression model and a classification model, specifically in the context of machine learning.
Regression models have response values that are quantitative while classification models have qualitative values. 

## Question 3:
#### Name two commonly used metrics for regression ML problems. Name two commonly used metrics for classification ML problems.
Two commonly used metrics for regression ML problems are Mean Squared Error and Mean Absolute Error. Two commonly used metrics for classification ML problems are Accuracy and Confusion Matrix. 

## Question 4:
#### As discussed, statistical models can be used for different purposes. These purposes can generally be classified into the following three categories. Provide a brief description of each.

##### Descriptive models:
Descriptive models seek to visually emphasize trends in the data. Typically are histograms, pie charts, bars or scatter plots.

##### Inferential models:
Inferential models investigate relationships between outcomes and predictors. Also, they assess whether trends contained in the research sample can be generalized to the larger populations.

##### Predictive models:
Predictive models are used to predict future events or outcomes by analyzing patterns in a given set of input data with minimum reducible error.

## Question 5:
#### Predictive models are frequently used in machine learning, and they can usually be described as either mechanistic or empirically-driven. Answer the following questions.
##### Define mechanistic. Define empirically-driven. How do these model types differ? How are they similar?
A mechanistic model predicts future values based on an assumed parametric form of “f”. An empirically-driven model predicts the future based on evidence, direct correlation between inputs and outputs. The models differ because empirically-driven models make no assumption about “f” and are similar because they both predict the future and can be overfit.

##### In general, is a mechanistic or empirically-driven model easier to understand? Explain your choice.
Mechanistic modeling is easier because it is based on assumptions rather than experimenting.

##### Describe how the bias-variance tradeoff is related to the use of mechanistic or empirically-driven models.
Bias-variance tradeoff relates to empirically-driven models because when using experimentally data variance can be high and needs to be minimized thus to minimize variance, bias must be increased. 

## Question 6:
#### A political candidate’s campaign has collected some detailed voter history data from their constituents. The campaign is interested in two questions:
##### - Given a voter’s profile/data, how likely is it that they will vote in favor of the candidate?
##### - How would a voter’s likelihood of support for the candidate change if they had personal contact with the candidate?
#### Classify each question as either predictive or inferential. Explain your reasoning for each.
“Given a voter’s profile/data, how likely is it that they will vote in favor of the candidate?” is a predictive question. This is because the probability of each candidate winning is being predicted. 
“How would a voter’s likelihood of support for the candidate change if they had personal contact with the candidate?” is an inferential question. This is because the question is testing if this event has an influence on the outcome. 

# Exploratory Data Analysis
## Exercise 1:
#### We are interested in highway miles per gallon, or the hwy variable. Create a histogram of this variable. Describe what you see/learn.
```{r}
hwy <- mpg$hwy
hist(hwy)
```

From the histogram, it seems that most cars have 15-10 or 25-30 highway miles per gallon. 

## Exercise 2:
#### Create a scatterplot. Put hwy on the x-axis and cty on the y-axis. Describe what you notice. Is there a relationship between hwy and cty? What does this mean?
```{r}
cty <- mpg$cty
plot(hwy, cty, main = "Scatterplot", xlab = "Highway Miles Per Gallon", ylab = "City Miles Per Gallon")
```

There seems to be a relationship between hwy and cty because an increasing linear line can be fit to the graph. I believe that it means that if the car has a high hwy it would also have a high cty and vice-versa. 

## Exercise 3:
#### Make a bar plot of manufacturer. Flip it so that the manufacturers are on the y-axis. Order the bars by height. Which manufacturer produced the most cars? Which produced the least?
```{r}
manu <- mpg$manufacturer
counts <- table(manu)
barplot(counts, horiz = TRUE, cex.names = 0.4)
```

Dodge produced the most cars while Lincoln produced the least.

## Exercise 4:
#### Make a box plot of hwy, grouped by cyl. Do you see a pattern? If so, what?
```{r}
boxplot(hwy ~ mpg$cyl)
```

Cars with more cylinders typically have lower highway miles per gallon. 

## Exercise 5:
#### Which variables are positively or negatively correlated with which others? Do these relationships make sense to you? Are there any that surprise you?
```{r}
library(corrplot)
num.mpg <- mpg %>% select_if(is.numeric)
M <- cor(num.mpg)
corrplot(M, type = "lower")
```

The variables in red are negatively correlated while the ones in blue are positively. These relationships make sense because, for example, if a car has a high highway mpg it would also have a relatively high city mpg. The data regarding cyl and displ surprise me because I do not know much about them. 

# 231 Only 
## Exercise 6:
```{r}
library(ggplot2)
library(ggthemes)
ggplot(mpg, aes(x=hwy, y=class)) + 
  geom_boxplot(color = "black") +
  theme_minimal() +
  theme(axis.line.x = element_line(colour = 'black', size=0.5, linetype='solid')) +
  labs(x="Highway MPG", y = "Vehicle Class") +
  geom_jitter(color = "gray")
```

## Exercise 7:
```{r}
ggplot(mpg, aes(y=hwy, x=class, fill= drv)) + 
  geom_boxplot()
```

## Exercise 8:
```{r}
ggplot(mpg, aes(y=hwy, x=displ, fill= drv)) + 
  geom_point(aes(colour = drv)) +
  geom_smooth(se = FALSE, aes(linetype=drv))
```
